### 2018-10-11 Simulator/article.tex
Testing change of frame letters

\newif\ifFlatArchive\FlatArchivefalse           % Whether archive is flat (messy) or contain 
                                                % subfolders for graphics etc.
\newif\ifFloatAtEnd\FloatAtEndfalse             % Available in PeerReview mode:
                                                % Place floats at end of document?
\newif\ifTODO\TODOtrue                        % Use todo notes?

\newif\ifBuildBibliography\BuildBibliographyfalse  % Whether to include the references.bib file or use inline references.

@@

\usepackage{cite}

@@

\usepackage[cmex10]{amsmath}                    % cmex10 option to be IEEE explore compliant
\interdisplaylinepenalty=2500                   % Allows multiline equations to be broken

% \RequirePackage{amssymb}

@@

@@

% \ifPeerReview
% \let\OldIncludegraphics{\includegraphics}
% \usepackage{letltxmacro}
% \LetLtxMacro{\OldIncludegrsaphics}{\includegraphics}
% \renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=\linewidth, #1]{#2}}
% \fi

@@

% \usepackage[maxfloats=40]{morefloats}
\newcounter{todoidx}
% \setcounter{todoidx}

\ifTODO
   \definecolor{todobackground}{rgb}{0.95,0.95,0.95}
   \setlength\marginparsep{1pt}
   \setlength\marginparwidth{35pt}
   \newlength\marginparwidthsmall
   \setlength\marginparwidthsmall{\marginparwidth}
   \addtolength\marginparwidthsmall{-7pt}
   \newcommand\todo[1]{%
      \addtocounter{todoidx}{1}%
      {\color{Red}\bf(\thetodoidx{})}%%\fbox{\bf\thetodoidx{}}}%
      \marginpar{%
         {\vspace*{-10pt}\color{Red}\fbox{\bf\thetodoidx{}}}\\%
         \fcolorbox{red}{todobackground}{\parbox{\marginparwidthsmall}{\raggedright\scriptsize #1}}}}

   \newcommand\todopar[1]{\fcolorbox{red}{white}{\parbox{0.97\linewidth}{#1}}}
\else
%    \usepackage[disable]{./todonotes} 
   \newcommand\todo[1]{}
\fi

>>>>

\newif\ifFlatArchive\FlatArchivefalse           % Whether archive is flat (messy) or contain 
                                                % subfolders for graphics etc.
\newif\ifFloatAtEnd\FloatAtEndfalse             % Available in PeerReview mode:
                                                % Place floats at end of document?
\newif\ifTODO\TODOtrue                        % Use todo notes?
\newif\ifOPTtodos\OPTtodosfalse
\newif\ifPhdDoc\PhdDocfalse

\newif\ifBuildBibliography\BuildBibliographytrue  % Whether to include the references.bib file or use inline references.
\newif\ifOverLeaf\OverLeaftrue                  % Using OverLeaf to edit?

@@

%\usepackage[backend=biber,bibstyle=ieee]{biblatex}
%\addbibresource{references.bib}

\RequirePackage{amsfonts}
\usepackage{cite}

@@

%\usepackage[cmex10]{amsmath}                    % cmex10 option to be IEEE explore compliant
\usepackage[cmex10]{mathtools}                    % cmex10 option to be IEEE explore compliant
\interdisplaylinepenalty=2500                   % Allows multiline equations to be broken
\RequirePackage{amssymb}

@@

\usepackage{accents}

@@

\usepackage{tikz}
\usetikzlibrary{calc,positioning,shapes,snakes}
%\newcommand{\tikzmark}[1]{\tikz[remember picture]\coordinate(#1){(-.35em,-.545ex)};}
\newcommand{\tikzmark}[2]{\tikz[remember picture,baseline=(#1.base)]\node[shape=rectangle,inner sep=0pt](#1){#2};}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\usepackage{multirow}
\usepackage{tabularx}

%\usepackage[active,float]{preview}
%\PreviewEnvironment{gather} 
% \ifPeerReview
% \let\OldIncludegraphics{\includegraphics}
% \usepackage{letltxmacro}
% \LetLtxMacro{\OldIncludegrsaphics}{\includegraphics}
% \renewcommand{\includegraphics}[2][]{\OldIncludegraphics[width=\linewidth, #1]{#2}}
% \fi

@@

% \usepackage[maxfloats=40]{morefloats}
%\newcounter{todoidx}
%% \setcounter{todoidx}
%
%\ifTODO
%   \definecolor{todobackground}{rgb}{0.95,0.95,0.95}
%   \setlength\marginparsep{1pt}
%   \setlength\marginparwidth{35pt}
%   \newlength\marginparwidthsmall
%   \setlength\marginparwidthsmall{\marginparwidth}
%   \addtolength\marginparwidthsmall{-7pt}
%   \newcommand\todo[1]{%
%      \addtocounter{todoidx}{1}%
%      {\color{Red}\bf(\thetodoidx{})}%%\fbox{\bf\thetodoidx{}}}%
%      \marginpar{%
%         {\vspace*{-10pt}\color{Red}\fbox{\bf\thetodoidx{}}}\\%
%         \fcolorbox{red}{todobackground}{\parbox{\marginparwidthsmall}{\raggedright\scriptsize #1}}}}
%
%   \newcommand\todopar[1]{\fcolorbox{red}{white}{\parbox{0.97\linewidth}{#1}}}
%\else
%%    \usepackage[disable]{./todonotes} 
%   \newcommand\todo[1]{}
%\fi

\newcommand\gls[1]{}

### 2018-10-11 ThesisJo/2_publications.tex
Almost perfect build of all targets with precompiled header
Almost there with all contents, but encoding issues. Trying Xelatex
Testing change of frame letters

%&2_publications
% !TeX root = 2_publications

\input{header}
\input{subheader}
\endofdump

\input{glossary}
\makeglossaries

\begin{document}
%\documentclass[float=false, crop=false]{standalone}
%\usepackage[subpreambles=true]{standalone}
%\input{header}
%\input{subheader}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper I}\label{sec:paperI} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\paper{I} describes the use and implementation of the \gls{MVDR} beamformer in active sonar imaging in its most general and exact form.

Its widespread use and generic implementation makes it the reference. thus will be considered the reference method. , and this will be referred to as a fully adaptive technique. Due to its ability to analytically  to apply the minimum variance criteria in an exact manner, Due to its popularity and We view this as the reference method for We refer to it as fully adaptive since it computes analytically evaluates   yields  This is a fully adaptive method   This method is  on a \gls{GPGPU} for active sonar imaging. This is to prove that using a \gls{MVDR}

Paper II takes the \gls{MVDR} method and applies it to clinical Ultrasound Imaging

We point out important hardware limitations for these devices, and assess the design in terms of how efficiently it is able to use the GPU's resources. On a quad-core Intel Xeon system with a high-end Nvidia GPU, our GPU implementation renders more than a million pixels per second (1\;MP/s). Compared to our initial central processing unit (CPU) implementation the optimizations described herein led to a speedup of more than two orders of magnitude, or an expected five to ten times improvement had the CPU received similar optimization effort. This throu

@@

\begin{itemize}
\item Minimum variance
\item Full search space, continuous
\item Element-space
\item Sonar
\item Computationally complex
\item Common: Minimal temporal statistics (near single snapshot)
\end{itemize}

This is a popular method that analytically estimates the array's weight set that best comply with the minimum variance criterion. Its analytic nature allows it to chose from an infinite number of possible weight sets, but with a solution space limited by the constraint of unit gain in the look direction.

The first paper is concerned with applying the Minimum Variance Distortionless Response (MVDR) beamformer to active sonar imaging, with emphasis on performing arithmetic reductions and solving the practical its high computational complexity. It shows that  with emphasis on solving and with the process of implementing this method on a \gls{GPU}. The goal is to improve sonar image quality without suffering the consequences of the MVDR's inherently high computational complexity.

Combined, these steps led to a speedup of an order of magnitude or more compared to its straight-forward \gls{CPU} implementation.

In functionality this method acts as our reference method, but its practicality is limited by a high computational complexity. The paper suggests means for reducing the arithmetic complexity and provide extensive details for fitting the implementation on a \gls{GPU} for a major speedup. It also relates the performance to theoretical maximums and investigates remaining bottlenecks.

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper III}\label{sec:paperIII} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{Low Complexity Adaptive Beamforming}~\cite{Buskenes2015}\\
J.I. Buskenes, R.E. Hansen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 42, Issue 1, January 2017

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper IV}\label{sec:paperIV} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{itemize}
\item Template-generation
\item Autonomy
\item Sonar
\end{itemize}

@@

>>>>

%&header
% !TeX root = 2_publications
\input{subheader}
\endofdump

\ifRootBuild\else
  \input{glossary}
  \makeglossaries
\fi

\begin{document}

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Intro}\label{sec:intro} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%
%\paper{I} describes the use and implementation of the \gls{MVDR} beamformer in active sonar imaging in its most general and exact form. 
%
%Its widespread use and generic implementation makes it the reference. thus will be considered the reference method. , and this will be referred to as a fully adaptive technique. Due to its ability to analytically  to apply the minimum variance criteria in an exact manner, Due to its popularity and We view this as the reference method for We refer to it as fully adaptive since it computes analytically evaluates   yields  This is a fully adaptive method   This method is  on a \gls{GPGPU} for active sonar imaging. This is to prove that using a \gls{MVDR}
%
%Paper II takes the \gls{MVDR} method and applies it to clinical Ultrasound Imaging
%
%
%We point out important hardware limitations for these devices, and assess the design in terms of how efficiently it is able to use the GPU's resources. On a quad-core Intel Xeon system with a high-end Nvidia GPU, our GPU implementation renders more than a million pixels per second (1\;MP/s). Compared to our initial central processing unit (CPU) implementation the optimizations described herein led to a speedup of more than two orders of magnitude, or an expected five to ten times improvement had the CPU received similar optimization effort. This throu

Common:
\begin{itemize}\setlength\baselineskip{0pt}
\item Adaptive
\item Processing speed
\end{itemize}

@@

\begin{itemize}\setlength\baselineskip{0pt}
\item Optimization: Minimum variance
\item Solution space: Continuous, closed when made robust
\item Time domain
\item Sonar
\item Computationally complex, $O(N^3)$.
\item Information significance: Minimal temporal sample support (~single snapshot)
\end{itemize}

%This is a popular method that analytically estimates the array's weight set that best comply with the minimum variance criterion. Its analytic nature allows it to chose from an infinite number of possible weight sets, but with a solution space limited by the constraint of unit gain in the look direction. 
%
%The first paper is concerned with applying the Minimum Variance Distortionless Response (MVDR) beamformer to active sonar imaging, with emphasis on performing arithmetic reductions and solving the practical its high computational complexity. It shows that  with emphasis on solving and with the process of implementing this method on a \gls{GPU}. The goal is to improve sonar image quality without suffering the consequences of the MVDR's inherently high computational complexity.
%
%Combined, these steps led to a speedup of an order of magnitude or more compared to its straight-forward \gls{CPU} implementation.
%
%In functionality this method acts as our reference method, but its practicality is limited by a high computational complexity. The paper suggests means for reducing the arithmetic complexity and provide extensive details for fitting the implementation on a \gls{GPU} for a major speedup. It also relates the performance to theoretical maximums and investigates remaining bottlenecks.

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper III}\label{sec:paperIII} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{Low Complexity Adaptive Beamforming}~\cite{Buskenes2014}\\
J.I. Buskenes, R.E. Hansen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 42, Issue 1, January 2017

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper IV}\label{sec:paperIV} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{A Real-Time Synthetic Aperture Sonar Simulator for Automatic Target Recognition: Notation and application}\\
J.I. Buskenes, H. Midelfart\\
\textit{IEEE Journal of Oceanic Engineering}\\
Submitted DD.MM.2018%\\\\
%Volume 42, Issue 1, January 2017

\begin{itemize}
\item Template-generation
\item Autonomy
\item Sonar
	\item Incredibly fast template matching possible.
	\item Possibly a powerful tool to pass prior knowledge to deep learning
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
\end{itemize}

@@

Machine cognition:
\begin{itemize}
\item Ethical dilemmas. Decision and actions have consequences. Who's legally responsible for a machine's doing? Itself?
\item Safety. Machine flexibility allow more possible outcomes, and a wider range of security issues to consider.
\end{itemize}

### 2018-10-11 ThesisJo/3_discussion.tex
Almost there with all contents, but encoding issues. Trying Xelatex
Testing change of frame letters

%&header
% !TeX root = 3_discussion
\documentclass[thesis.tex]{subfiles}
\input{subheader}
\begin{document}

@@

\begin{itemize}
\item Improving imaging systems by making them smarter.
\item Big data - more processing power.
\item Processing speed
\begin{itemize}
	\item Choose the best tool (algorithm) for the job
	\item More complex not always better, but slower
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
\end{itemize}
\item Faster IS smarter
 \begin{itemize}
 	\item Short feedback loops allow real-time adaptivity
 	\item More data contains more information
 	\item Faster learning
 \end{itemize}
\item Smarter IS better
 \begin{itemize}
	\item Better image quality
	\item Improved autonomy
\end{itemize}
\end{itemize}

\begin{itemize}
\item Improving imaging systems by making them smarter.
\item Big data - more processing power.
\item IV - Simulator
\begin{itemize}
	\item Incredibly fast template matching possible.
	\item Possibly a powerful tool to pass prior knowledge to deep learning
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
\end{itemize}
\item Faster IS smarter
 \begin{itemize}
 	\item Short feedback loops allow real-time adaptivity
 	\item More data contains more information
 	\item Faster learning
 \end{itemize}
\item Smarter IS better
 \begin{itemize}
	\item Better image quality
	\item Improved autonomy
\end{itemize}
\end{itemize}

>>>>

%&header
% !TeX root = 3_discussion
\input{subheader}
\endofdump

\ifRootBuild\else
  \input{glossary}
  \makeglossaries
\fi

\begin{document}

@@

Constrain attention to beamformers and simulators (not ATR?).

\section{Trends}

The future is faster, smarter.

\subsection{Hardware}

\begin{itemize}
\item Ever increasing processing speed.
\begin{itemize}
	\item Choose the best tool (algorithm) for the job
	\item More complex not always better, but slower
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
   \item Intel entering the discrete graphics stage
   \end{itemize}
\item Software defined transceiver/receivers
\begin{itemize}
\item System on Chips with integrated ADCs/DACs. Big power in small package.
\item Compact, flexible, direct conversion
\item Software defined signal generation, filtering, synchronization, rate-conversion and mixing.
\item Multipurpose, hardware reuse, easier implementation of advanced signal processing, faster development and lower price.
\item 
\item Compact multi-band Wifi
\item MIMO (in new wifi). Breaks array SNR boost. Not viable for long range.
\item Phones: Wifi 2.4-2.5G, 5G (massive MIMO, heavy SDR, 600M-6G), LTE (4G, Long Term Evolution 800M), GSM (2G, Global System for Mobile communications 900M,1.8-1.9M), UMTS (3G, Universal Mobile Telecommunications System 1.9-2.1M), GPS (Global Positioning System 1.5G), Bluetooth (2.4G), NFC (Near-Field Communications 13.56M), FM (100M). Multi-radio in a single SDR package!
\end{itemize}
\item Short feedback loops allow real-time adaptivity
\end{itemize}

\subsection{Algorithms}

\begin{itemize}
\item Intelligence - Ability to process data fast, and retain key information
\item Adaptive methods ever more attractive.
   \begin{itemize}
   \item Processing more data to increase significance of extracted information
   \item Reducing reaction time: Act when the data is fresh/valid.
   \end{itemize}
\item Better image quality
\item Improved autonomy
\end{itemize}

\subsection{Connectivity}

\begin{itemize}
\item Internet of Things
\item Big data
\item Intelligent mesh of nodes
\end{itemize}

\section{Risks \& Challenges}

\begin{itemize}
\item Improving imaging systems by making them smarter.
\item Big data - more processing power.
\item Faster IS smarter
 \begin{itemize}
 	\item Short feedback loops allow real-time adaptivity
 	\item More data contains more information
 	\item Faster learning
 \end{itemize}
\end{itemize}

%Discussion and future work
%   Summary and discussion
%      Simulator:
%      - The way to train and constrain deep learning algorithms?
%      MVDR:
%      - Analytical optimization - full adaptivity.
%      - Heavy on computations
%      LCA:
%      - Optimization by trial and error - discrete solution space.
%      - Works almost as well as MVDR in most cases
%      - Fast!
%      - Well suited for hardware
%      
%      Lessions from deep learning:
%      - 
%      
%   Future work

\section{MVDR}

Recent developments:
- A low complexity minimum variance beamformer for ultrasound imaging using dominant mode rejection~\cite{Jiang2016}
- Google deep beamforming

\begin{itemize}
\item SDR - Entirely software defined radio, radar and sonar, bigger potential for adaptive methods
\item Cognitive radar/sonar. 1. Adaptive signal processing, learn from data. 2. Feedback from receiver to transmitter. 3. Preservation of the information in received data. All of this is present in bats.
\end{itemize}

### 2018-10-11 ThesisJo/frontmatter.tex
Almost there with all contents, but encoding issues. Trying Xelatex
Testing change of frame letters

\input{variables}\ifMonolithic\else\input{header}\input{subheader}\fi
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

@@

\newpage
\chapter*{Related Publications}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\ifMonolithic\else\input{subfooter}\fi

>>>>

%&header
% !TeX root = thesis
\input{subheader}
\endofdump

\ifRootBuild\else
  \input{glossary}
  \makeglossaries
\fi

\begin{document}

\begin{markdown}
Hey
\end{markdown}

@@

\newpage
\chapter*{Abstract}

@@

This thesis is based on the following four journal papers, referred to by their roman literal:

\newcommand\pubEntry[4]{\item #2, "#1", \textit{#3}, #4}
{\renewcommand{\theenumi}{\textbf{\Roman{enumi}}}
\renewcommand\cleardoublepage{\newpage}
\begin{enumerate}
\pubEntry{%
An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 40, Issue 2, April 2015}
%
\pubEntry{%
Implementing Capon Beamformer on a GPU for Real-Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng and S. Holm}%
{IEEE Transactions on Ultrasonics, Ferroelectrics, and Frequency Control}%
{Volume 61, Issue 1, January 2015}
%
\pubEntry%
{Low Complexity Adaptive Beamforming}%
{J.I. Buskenes, R.E. Hansen and A. Austeng}%
{IEEE Journal of Oceanic Engineering}%
{Volume 42, Issue 1, January 2017}
%
\pubEntry%
{A Real-Time Synthetic Aperture Sonar Simulator for Automatic Target Recognition: Notation and application}%
{J.I. Buskenes, H. Midelfart}%
{IEEE Journal of Oceanic Engineering}%
{Submitted DD.MM.2018}
\end{enumerate}
%}

%\newpage
\chapter*{Related Publications}

\begin{enumerate}
\pubEntry{%
A GPU sonar simulator for automatic target recognition}%
{J.I. Buskenes, H. Midelfart, Ø. Midtgaard}%
{Underwater acoustics international conference and exhibition}%
{June 2014}
%
\pubEntry{%
Adapting the minimum variance beamformer to a graphics processing unit for active sonar imaging systems}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{Journal of the Acoustical Society of America}%
{Volume 133, 2013}
%
\pubEntry%
{GPU-Based Adaptive Beamforming for Active Sonar Imaging}%
{J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen, and A. Austeng}%
{11th European Conference on Underwater Acoustics (ECUA)}%
{ISBN: 978-1-62276-192-0 p. 860-865, 2012}
%
\pubEntry%
{Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging}%
{J.P. Åsen, J.I. Buskenes, C.-I.C. Nilsen, A. Austeng, and S. Holm}%
{Proceedings - IEEE Ultrasonics Symposium}%
{p. 2133-2136, 2012}
%
\pubEntry%
{A Low Complexity Adaptive Beamformer for Active Sonar Imaging}%
{J.I. Buskenes, C.-I.C. Nilsen, and A. Austeng}%
{Underwater Acoustic Measurements: Technology and Results}%
{2011}
\end{enumerate}
}

\end{document}
